---
title: "Clinton and Eisenhower Speeches"
author: "Janie Briglio and Kerin Grewal"
date: "April 8, 2018"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(readtext)
library(dplyr)
library(tidytext)
library(ggplot2)
library(tidyr)
library(wordcloud)
library(reshape2)
library(gridExtra)
clinton.1 <- readtext("clinton.txt") 
clinton.1 <- as.data.frame(clinton.1) %>% unnest_tokens(word, text)
data(stop_words)
clinton.2 <- clinton.1 %>% anti_join(stop_words) 
clinton.3 <- clinton.2 %>% count(word, sort = TRUE)
clinton.4 <- as_tibble(clinton.3)

clinton_top <- top_n(clinton.4, 10)

eisenhower.1 <- readtext("eisenhower.txt") 
eisenhower.1 <- as.data.frame(eisenhower.1) %>% unnest_tokens(word, text)
data(stop_words)
eisenhower.2 <- eisenhower.1 %>% anti_join(stop_words) 
eisenhower.3 <- eisenhower.2 %>% count(word, sort = TRUE)
eisenhower.4 <- as_tibble(eisenhower.3)

eisenhower_top <- top_n(eisenhower.4, 10)

```


##The most used words in each speech
```{r, echo=FALSE}
plot1 <- clinton_top %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  ylim(0, 50) +
  geom_col() +
  labs(title = "Clinton") +
  xlab(NULL) +
  coord_flip()

plot2 <- eisenhower_top %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  ylim(0, 50) +
  geom_col() +
  labs(title= "Eisenhower") +
  xlab(NULL) +
  coord_flip()

grid.arrange(plot2, plot1)
```

```{r, include= FALSE}
clinton.sentiment <- clinton.4 %>%
  inner_join(get_sentiments("bing")) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

eisenhower.sentiment <- eisenhower.4 %>%
  inner_join(get_sentiments("bing")) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

clinton.sentiment2 <- clinton.4 %>%
  inner_join(get_sentiments("nrc")) %>%
  spread(sentiment, n, fill = 0) %>% 
  select("word", "joy", "anger", "anticipation")

eisenhower.sentiment2 <- eisenhower.4 %>%
  inner_join(get_sentiments("nrc")) %>%
  spread(sentiment, n, fill = 0) %>% 
  select("word", "joy", "anger", "anticipation")

clinton.sentiment3 <- clinton.4 %>%
  inner_join(get_sentiments("afinn"))

eisenhower.sentiment3 <- eisenhower.4 %>%
  inner_join(get_sentiments("afinn"))
```

###President Dwight D. Eisenhower delivered his state of the union address on Thursday, January 6, 1955. This was just after WWII ended and FDR left office. His speech addressed how to avoid a "nuclear holocaust", for at the time there appeared to be a high liklihood of nuclear warfare. He also discussed the "anxieties of this divided world". On a more positive note, he mentioned that the country's economic sucesses could be attributed to raw materials and energy, and how to continue to have such successes without exploitation. 

###President Bill Clinton delivered his first state of the union address on January 25th, 1994. In his speech he promoted the Clinton health care plan, wanted to guarentee every American access to private health insurance. He also wanted to cut down on crime and expand gun control and access to drug treatment and education. Clinton's presidency was during the longest period of peacetime economic expansion in American history. He also left office with the highest end-of-office approval rating of any U.S. preisdent since WWII.

\pagebreak


##Comparing positive and negative sentiments in each speech
###In Clinton's speech, the beginning of the speech has a high concentration of very negative sentients, then the speech ends with a high concentration in more positive sentiments. While Eisenhower's speech also begins with more negative sentiments, they are not as extreme. Also, there is a large spike immediately after that in postitive sentiments then it levels off to less emotionally charged language. 
```{r, echo = FALSE}
clinton.plot <- ggplot(clinton.sentiment, aes(word, sentiment, fill = word)) +
  geom_col(show.legend = FALSE)+
  ylim(-17, 17) +
  labs(title = "Clinton") +
  theme(axis.text.x=element_blank(),
      axis.ticks.x=element_blank())

eisenhower.plot <- ggplot(eisenhower.sentiment, aes(word, sentiment, fill = word)) +
  geom_col(show.legend = FALSE)+
  ylim(-17, 17) +
  labs(title = "Eisenhower") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

eisenhower.plot
clinton.plot
```

\pagebreak 

##Comparing presence of joy, anger and anticipation in speeches
###Clinton's Speech 
###Clinton's speech uses many "joy" words through out the speech, with a high concentration of "anger" words towards the beginning and then ends the speech with "anticipation" words. 
```{r, echo = FALSE}
ggplot(clinton.sentiment2, aes(word, joy, fill = word)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ggplot(clinton.sentiment2, aes(word, anger, fill = word)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ggplot(clinton.sentiment2, aes(word, anticipation, fill = word)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

\pagebreak

###Here are the top 10 joyful and angry words in Clinton's speech, represented in the graphs above.
```{r, echo=FALSE}
clinton.joy <- subset(clinton.sentiment2, select = c(word, joy))
clinton.joy <- clinton.joy[-row(clinton.joy)[clinton.joy == 0],]
clinton.joy1 <- clinton.joy[order(-clinton.joy$joy), ]

clinton.joy1 <- head(clinton.joy1, 10)

joy <- clinton.joy1 %>%
  mutate(word = reorder(word, joy)) %>%
  ggplot(aes(word, joy)) +
  geom_col() +
  ylim(0, 25) +
  xlab(NULL) +
  labs(title = "Joy") +
  coord_flip()

clinton.anger <- subset(clinton.sentiment2, select = c(word, anger))
clinton.anger <- clinton.anger[-row(clinton.anger)[clinton.anger == 0],]
clinton.anger1 <- clinton.anger[order(-clinton.anger$anger), ]

clinton.anger1 <- head(clinton.anger1, 10)

anger <- clinton.anger1 %>%
  mutate(word = reorder(word, anger)) %>%
  ggplot(aes(word, anger)) +
  geom_col() +
  ylim(0, 25) +
  xlab(NULL) +
  labs(title = "Anger") +
  coord_flip()

grid.arrange(joy, anger)
```

####*It is important to note that money is considered both an angry and joyful word, as it can be interperted as either of the two, situationally.

\pagebreak

###Eisenhower's Speech 
###Eisenhower's speech begins with more "anger" words, then transitions into more "joy" words, which continues until the end of the speech. Close to the middle of the speech, "anticipation" words peak and continue until the end of the speech. 
```{r, echo=FALSE}
ggplot(eisenhower.sentiment2, aes(word, joy, fill = word)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ggplot(eisenhower.sentiment2, aes(word, anger, fill = word)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ggplot(eisenhower.sentiment2, aes(word, anticipation, fill = word)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

\pagebreak

###Here are the top 10 joyful and angry words in Eisenhower's speech, represented in the graphs above.
```{r, echo=FALSE}
eisenhower.joy <- subset(eisenhower.sentiment2, select = c(word, joy))
eisenhower.joy <- eisenhower.joy[-row(eisenhower.joy)[eisenhower.joy == 0],]
eisenhower.joy1 <- eisenhower.joy[order(-eisenhower.joy$joy), ]

eisenhower.joy1 <- head(eisenhower.joy1, 10)

joy1 <- eisenhower.joy1 %>%
  mutate(word = reorder(word, joy)) %>%
  ggplot(aes(word, joy)) +
  ylim(0, 25) +
  geom_col() +
  xlab(NULL) +
  labs(title = "Joy") +
  coord_flip()

eisenhower.anger <- subset(eisenhower.sentiment2, select = c(word, anger))
eisenhower.anger <- eisenhower.anger[-row(eisenhower.anger)[eisenhower.anger == 0],]
eisenhower.anger1 <- eisenhower.anger[order(-eisenhower.anger$anger), ]

eisenhower.anger1 <- head(eisenhower.anger1, 10)

anger1 <- eisenhower.anger1 %>%
  mutate(word = reorder(word, anger)) %>%
  ggplot(aes(word, anger)) +
  ylim(0, 25) +
  geom_col() +
  xlab(NULL) +
  labs(title = "Anger") +
  coord_flip()

grid.arrange(joy1, anger1)
```


\pagebreak 

##Comparing words by AFINN score

###Both of the speeches have very similar AFINN scores throughout their speeches. The scores for both are fairly balances between positive and negative.
```{r, echo=FALSE}
ggplot(clinton.sentiment3, aes(word, score, fill = word)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Clinton") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
ggplot(eisenhower.sentiment3, aes(word, score, fill = word)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Eisenhower") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```